From Real-time Attention Assessment to ``With-me-ness'' in Human-Robot Interaction

### 2015-09-18 paper.tex
Done with a first pass on section 2

Besides being fast, the face alignment algorithm has been found to perform well,
including in a range of difficult situations (occlusions, head angle) as
encountered in field experiments.  Figure~\ref{head_pose_real_world} shows a
selection of such difficult scenes on one child.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\linewidth]{head_pose}
    \caption{The 6D head pose is estimated by fitting a 3D model of an
        adult head (left) onto the detected 2D features of the face (right). We
        rely on an iterative $PnP$ algorithm, using 8 correspondence pairs
        (three are depicted: the sellion -- the nasal depression --, the left
        tragion and the menton).}

@@

\subsection{Field of Attention}

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{field_of_attention}
    \caption{\small The visual field of attention is approximated to a
        20$\degree$ cone, spanning from the head's sellion. The objects whose 3D
    pose intersect with this cone are considered \emph{in focus}.}
    \label{rviz}
\end{figure}

The third step is to generate a representation of the field of view whose origin
and orientation are the ones defined by the face detection \textit{tf} frame.

The breadth of the human field of attention has been discussed in the
literature.
Holmqvist in~\cite{holmqvist2011eye} specifies that the visual human
range\fixme{field of view or field of attention?} is $
\pm  40\degree $ in the horizontal and $ \pm 25\degree $ in the vertical. In
\cite{walker1980clinical} Spector provides a more detailed specification for
each eye splitting the vertical range into $ 60\degree $ the upper region and $
75\degree $ the lower one. Moreover, the horizontal range gets separated in $
60\degree $ inwards (towards the nose) and $ 95\degree $ outwards. In robotics,
Sisbot\etal~\cite{sisbot2011situation} use a cone of 30$\degree$. In the
present implementation, the first approach explained has been chosen
representing the field of view using a cone with such dimensions\fixme{which
one? 40deg? 25deg?}.

Assuming that the \textit{tf} frames representing the robotic agent and the VFoA
candidates are static, need to be monitored. This task consists on evaluating
the intersection between the \textit{tf} frames and the field of view: If one
focus of attention is within the region defined by the field of view, we assume
the subject is looking at it. In order to solve such operation mathematically it
is necessary to compute the transformation matrix between the coordinate frames
to be able to locate the monitored frames \textit{A} in the subject's coordinate
system \textit{B} (see equation \ref{eq:transform}).

\begin{equation}
v' = B(A^{-1})v
\label{eq:transform}
\end{equation}

where \textit{v} is a point in \textit{A} that becomes v' in coordinate system
\textit{B}. However, in practice there are, at least, three different
\textit{tf} frames involved in the transformation (subject-camera, camera-base,
base-focus). Following the transformation result, the distance of the monitored
frame with respect to the x-axis face \textit{tf} frame, $ d_{x_{axis}} $,
representing the subject's face can be computed as simple as in equation
\ref{eq:pitagoras}. Always considering the x-axis as the one defining the face's
orientation and thus field of view's main axis.

\begin{equation}
d_{x_{axis}} = \sqrt{t_y^2 + t_z^2}
\label{eq:pitagoras}
\end{equation}

It is necessary to compare the distance acquired $ d_{x_{axis}} $ with the
radius of the field of view at the target \textit{tf} frame x-coordinate $
r_{fov} $ as shown in equation \ref{eq:fovtf}.

\begin{equation}
r_{fov} = tan\left(\frac{fov}{2}\right) \cdot t_x
\label{eq:fovtf}
\end{equation}

where \textit{fov} is the aperture angle of the field of view. Finally, if $
d_{x_{axis}}<r_{fov} $, the target is within the subject's field of view.

% Schema to clarify this?

\paragraph{Acquiring the Focus of Attention}

Explain that: other modules have to provide the sitation assessment (what
is where?) in terms of published frames, and a last module compute the
intersection of the VFoA and these frames\fixme{TBD}

@@

\subsection{System Implementation}

>>>>

Besides being fast, the face alignment algorithm has been found to perform well
in terms of robustness, including in a range of difficult situations
(occlusions, head angle) as encountered in field experiments.
Figure~\ref{head_pose_real_world} shows a selection of such difficult scenes with
one child.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\linewidth]{head_pose}
    \caption{The 6D head pose is estimated by fitting a 3D model of an
        adult head (left) onto the detected 2D features of the face (right). We
        rely on an iterative $PnP$ algorithm, using 8 correspondence pairs
        (three are depicted: the sellion -- the nasal depression --, the left
        tragion and the menton). The 3D origin of the head is set at the sellion.}

@@

\subsection{Field \& Focus of Attention}

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{field_of_attention}
    \caption{\small The visual field of attention is approximated to a
        20$\degree$ cone, spanning from the head's sellion. The objects whose 3D
    pose intersect with this cone are considered \emph{in focus}.}
    \label{fig:vfoa}
\end{figure}

We model the field of attention as the central region of the field of view.
The field of view itself is approximated to a cone spanned from the nasal
depression (sellion) of the human face.

The width of the human field of attention has been discussed in the literature.
Holmqvist~\cite{holmqvist2011eye} specifies that the visual human
range\fixme{field of view or field of attention?} is $ \pm  40\degree $ in the
horizontal and $ \pm 25\degree $ in the vertical.
In~\cite{walker1980clinical}, Spector provides a more detailed specification
for each eye splitting the vertical range into $ 60\degree $ the upper region
and $ 75\degree $ the lower one. Moreover, the horizontal range gets separated
in $ 60\degree $ inwards (towards the nose) and $ 95\degree $ outwards.
Previous work on visual perspective taking for social robotics
application~\cite{sisbot2011situation} propose to use a cone of 30$\degree$ of
aperture. In the present implementation, the first approach explained has been
chosen representing the field of view using a cone with such
dimensions\fixme{which one? 40deg? 25deg?}.

We then approximate the visual focus of attention (VFoA) of the human to the
objects which lie inside the field of attention (Figure~\ref{fig:vfoa}). At a
given time, more than one object can therefore be \emph{in focus}.

Our implementation has two limitations: objects are approximated to their origin
(\ie objects are either in or out of the field of attention), and we do not
check actual visibility: one object could be hidden by another, it would still
be considered as in focus. We did not address these limitations since our
experimental setup (involving relatively small objects with no occlusions) did
not require it. Techniques for a more complete assessment of the visual
perspective can be found in~\cite{sisbot2011situation} for instance.

Within these limitations, computing if object $A(x_A,y_A,z_A)$ is in the field
of attention of the human requires first to transform the coordinates
$A(X_A,Y_A,Z_A)$ into the frame of the face, and then to verify the following
inequality (with $fov$ the aperture, and assuming that the main axis of the
field of attention is aligned on the face's $\vec{x}$):

\begin{align}
    \sqrt{Y_A^2 + Z_A^2} < tan\left(\frac{fov}{2}\right) \cdot X_A
\label{eq:fov}
\end{align}

Our approach assumes that the poses of the objects of interest are available to
the system: as described in section~\ref{sec:system}, our implementation relies
on the ROS {\sc tf} framework to manage and make available to all software
modules the list of poses of existing objects (represented as {\it frames}), and
dedicated perception modules are in charged of publishing up-to-date
informations regarding the location of the points of interest (the so-called
\emph{situation assessment}). Due to the nature of the experiment, most of the
points of interest considered for the experimental validation presented
hereafter are actually taken as static with respect to the robot.

@@

\subsection{System Implementation}
\label{sec:system}
